---
title: "Functional Purrrograming"
author: "Luke KLEIN"
date: "March 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functional Programming

Functions can be arguments, too.  

```{r}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)


col_median <- function(df) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- median(df[[i]])
  }
  output
}

col_mean <- function(df) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- mean(df[[i]])
  }
  output
}

col_sd <- function(df, fun) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- sd(df[[i]])
  }
  output
}


# All of the above can be achieved using the following function

col_summary <- function(df, fun) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- fun(df[[i]])
  }
  output
}

col_summary(df, fun = median)
col_summary(df, fun = mean)
col_summary(df, fun = sd)
```

## The `purrr` package

```{r}
# Load the purrr package
library(purrr)

# Use map_dbl() to find column means
map_dbl(df, mean)

# Use map_dbl() to column medians
map_dbl(df, median)

# Use map_dbl() to find column standard deviations
map_dbl(df, sd)

```

The map functions use the `...` ("dot dot dot") argument to pass along additional arguments to `.f` each time it’s called. For example, we can pass the `trim` argument to the `mean()` function.  Multiple arguments can be passed along using commas to separate them. For example, we can also pass the `na.rm` argument to `mean()`

```{r}
map_dbl(df, mean, trim = 0.5 , na.rm = TRUE)
```


# Shortcuts for `.f`


```{r}
cyl <- split(mtcars, mtcars$cyl)
str(cyl)
```

One can extract a desired dataframe in the usual `[[` subsetting manner.

```{r}
cyl[[1]]
```

