---
title: "Functional Purrrograming"
author: "Luke KLEIN"
date: "March 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functional Programming

Functions can be arguments, too.  

```{r}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)


col_median <- function(df) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- median(df[[i]])
  }
  output
}

col_mean <- function(df) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- mean(df[[i]])
  }
  output
}

col_sd <- function(df, fun) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- sd(df[[i]])
  }
  output
}


# All of the above can be achieved using the following function

col_summary <- function(df, fun) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- fun(df[[i]])
  }
  output
}

col_summary(df, fun = median)
col_summary(df, fun = mean)
col_summary(df, fun = sd)
```

## The `purrr` package

```{r}
# Load the purrr package
library(purrr)

# Use map_dbl() to find column means
map_dbl(df, mean)

# Use map_dbl() to column medians
map_dbl(df, median)

# Use map_dbl() to find column standard deviations
map_dbl(df, sd)

```

The map functions use the `...` ("dot dot dot") argument to pass along additional arguments to `.f` each time it’s called. For example, we can pass the `trim` argument to the `mean()` function.  Multiple arguments can be passed along using commas to separate them. For example, we can also pass the `na.rm` argument to `mean()`

```{r}
map_dbl(df, mean, trim = 0.5 , na.rm = TRUE)
```


# Shortcuts for `.f`


```{r}
cyl <- split(mtcars, mtcars$cyl)
str(cyl)
```

One can extract a desired dataframe in the usual `[[` subsetting manner.

```{r}
cyl[[1]]
```

Our goal is to fit a separate linear regression of miles per gallon (mpg) against weight (wt) for each group of cars in our list of data frames, where each data frame in our list represents a different group. How should we get started?

First, let's confirm the structure of this list of data frames. Then, we'll solve a simpler problem first: fit the regression to the first group of cars.

Examine the structure of cyl with str().
```{r}
str(cyl)
```

Extract the first data frame in cyl and assign it to the variable four_cyls.
```{r}
four_cyls <- cyl[[1]]
```

Fit a linear regression of miles per gallon on weight using four_cyls as the data argument to lm().
```{r}
lm( mpg ~ wt , data = four_cyls )
```

We now have a snippet of code that performs the operation we want on one data frame. One option would be to turn this into a function, for example:

```
fit_reg <- function(df) {
  lm(mpg ~ wt, data = df)
}
```

Then pass this function into map():

```{r}
fit_reg <- function(df) {
  lm(mpg ~ wt, data = df)
}

map(cyl, fit_reg)
```


But it seems a bit much to define a function for such a specific model when we only want to do this once. Instead of defining the function in the global environment, we will just use the function anonymously inside our call to map().

What does this mean? Instead of referring to our function by name in map(), we define it on the fly in the .f argument to map().

### Rewrite the map() call to use the anonymous function function(df) lm(mpg ~ wt, data = df)

```{r}
# Rewrite to call an anonymous function
map(cyl, function(df) lm(mpg ~ wt, data = df) )

map(cyl, ~ lm(mpg ~ wt , .))
```


## Pipes

purrr also includes a pipe operator: %>%. The pipe operator is another shortcut that saves typing, but also increases readability. The explanation of the pipe operator is quite simple: x %>% f(y) is another way of writing f(x, y). That is, the left hand side of the pipe, x, becomes the first argument to the function, f(), on the right hand side of the pipe.

Take a look at our code to get our list of models:

cyl <- split(mtcars, mtcars$cyl) 
map(cyl, ~ lm(mpg ~ wt, data = .))
We split the data frame mtcars and save it as the variable cyl. We then pass cyl as the first argument to map to fit the models. We could rewrite this using the pipe operator as:

split(mtcars, mtcars$cyl) %>% 
  map(~ lm(mpg ~ wt, data = .))
We read this as "split the data frame mtcars on cyl, then use map() on the result."

One of the powerful things about the pipe is we can chain together many operations. Here is our complete code, written with pipes, instead assigning each step to a variable and using it in the next step:

mtcars %>% 
  split(mtcars$cyl) %>%
  map(~ lm(mpg ~ wt, data = .)) %>%
  map(coef) %>% 
  map_dbl("wt")
We've written some code in the editor to pull out the R2 from each model. Rewrite the last two lines to use a pipe instead.

```{r}
# Define models (don't change)
models <- mtcars %>% 
  split(mtcars$cyl) %>%
  map(~ lm(mpg ~ wt, data = .))

# Rewrite to be a single command using pipes 
# Rewrite to be a single command using pipes 
summaries <- map(models, summary)
map_dbl(summaries, "r.squared")

# Becomes...

models %>% 
  map(summary) %>%
  map_dbl("r.squared")
```
