---
title: "Chapter 1"
author: "Luke KLEIN"
date: "March 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rewrite the call to follow best practices
`mean(0.1 , x=c(1:9, NA), TRUE)`

Beyond the first couple of arguments you should always use matching by name. It makes your code much easier for you and others to read. This is particularly important if the argument is optional, because it has a default. When overriding a default value, it's good practice to use the name.

Notice that when you call a function, you should place a space around `=` in function calls, and always put a space after a comma, not before (just like in regular English). Using whitespace makes it easier to skim the function for the important components.


```{r best-practice}
mean(c(1:9, NA), trim = 0.1, na.rm = TRUE)
```

# Scoping

Scoping describes where — not when — to look for a variable value.

When you call a function, a new working envirnment is created to execute the function's body.

# Data structures

`NULL` is the abscence of a vector, and has length 0.

'NA' is a *missing value* and has length 1.

## lists

Subsets with [,  [[, or $

* [ extracts sublists
* [[ and $ extract elements (i.e. remove a level of heirarchy from the list)

### subsetting lists

```{r}
a <- list(
  a = 1:3,
  b = "a string",
  c = pi,
  d = list(-1,-5),
  e = NULL
)

# Returns a list containing the d element
str(a[4])

# Return the d element itself
str(a[[4]])

str(a[[4]][1])

str(a[[4]][[1]])
```

## Exploring Lists

Often you won't know exactly what is inside a list. But, you'll need to figure it out to get some useful piece of data. Extracting elements from the output of the names() and str() functions is a great way to explore the structure of a list.

Calling names() on a list will give you names at the top level of the list and str() will give you a full description of the entire list (which can sometimes be a little overwhelming).

tricky_list has a regression model stored in it. Let's see if we can drill down and pull out the slope estimate corresponding to the wt variable.

```{r}
tricky_list = list(
  nums = rnorm(10),
  y = c(rep(FALSE,5),rep(TRUE,5)),
  x = list("hello","hi!","goodbye","bye!"),
  model = lm(formula = mpg ~ wt, data = mtcars)
)


# Guess where the regression model is stored
names(tricky_list)

# Use names() and str() on the model element
names(tricky_list[["model"]])
str(tricky_list[["model"]])

# Subset the coefficients element
tricky_list[["model"]]$coefficients

# Subset the wt element
tricky_list[["model"]][["coefficients"]][["wt"]]

# Try using the $ symbol
##  tricky_list$model$coefficients$wt  <-- causes an error
# Note that the coefficients variable is of type double, 
# therefor $ won't work on it.  However, [[ does
typeof(tricky_list$model$coefficients)

tricky_list$model$coefficients[["wt"]]
```





